The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-02-24 src/ms.tex
Start section introduction
Signed-off-by: nstarman <nstarkman@protonmail.com>

Summarize major pieces of development since the last paper (since v2.0). A few
ideas (taken from What's New pages) below, but feel free to expand on or remove
from this list:
\begin{itemize}
    \item Support for representing and transforming velocity data in coordinates, epoch propagation (v3.0)
    \item Improved support for astronomical time series: TimeSeries object (v3.2), Box Least Squares periodogram (v3.1)
    \item Overall improved support of Quantity throughout numpy (v4.0) and scipy
    \item Native support for Time, Quantity, and SkyCoord objects in Astropy tables
    \item APE 14 WCS
    \item A new SpectralCoord class for representing and transforming spectral quantities
    \item Support for dask/large arrays in several places in the core package (4.1, 5.0, and maybe other versions)
\end{itemize}

@@

All coordinated and affiliated packages are listed on the astropy
website\footnote{\url{https://www.astropy.org/affiliated}}. Since 
\cite{astropy:2018}, there has been an expansion of affiliated
packages for gravitational astrophysics including:
\texttt{PyCBC} for exploring gravitational wave signals, \texttt{lenstronomy} for
modeling strong gravitational lenses, \texttt{ligo.skymap} for visualizing
gravitational wave probability maps, and \texttt{EinsteinPy} for general
relativity and gravitational astronomy. There have been several packages
added to the ecosystem related to HEALPix: \texttt{astropy-healpix}
for a BSD-licensed HEALPix implementation, and \texttt{mocpy} for Multi-Order
Coverage maps. \texttt{astroalign} has been introduced for astrometric registration,
and \texttt{python-cpl} has been added for ESO pipelines and VLT data products. 
For ground-based astronomy, \texttt{baseband} has added IO capabilities for 
VLBI, and \texttt{SpectraPy} brings slit spectroscopy to the astropy ecosystem. 
We have \texttt{agnpy} for AGN jets, and \texttt{statmorph} for fitting galactic
morphological diagnostics. \texttt{saba} gives an interface to sherpa's fitting 
routines. For modeling interstellar dust extinction we have 
\texttt{dust_extinction}, and for extracting features from time series we 
have \texttt{feets}. We also have \texttt{corral} for managing data intensive parallel 
pipelines. \texttt{BayesicFitting} provides an interface for generic Bayesian 
inference, and \texttt{sbpy} \cite{Mommert2019} does calculations for asteroid and cometary
astrophysics. Finally, \texttt{synphot} provides an interface for synthetic photometry.

Several of the coordinated and affiliated packages described in \cite{astropy:2018} have had substantial improvements. \texttt{astroquery}
\cite{astroquery} has added
access to roughly a dozen new missions and data services, including JWST, and
the project has switched to a continuous release model. Every time a change
is committed to the main development branch it is published on the Python
Package Index (PyPI, ref) and available for installation. Formal releases are
still done a couple of times per year. \texttt{photutils} \cite
{photutils} released its first stable version, indicating that the API will
change less frequently, and there have been several significant performance
improvements.  \texttt{ccdproc} \cite{ccdproc} also released a new major
version, bringing better performance to some image combination operations.
The \texttt{regions} package \cite{pyregions}, for manipulating ds9-style
region definitions \cite{ds9}, added new ways to manipulate regions and
introduced new region types. \texttt{reproject}’s \cite{reproject} major new
feature is a function to align and co-add images to create a mosaic; better
support for parallelization was also added. \texttt{specutils}, the package
that defines containers for 1D and 2D spectra \cite{specutils}, also had its
first stable release and the addition of classes to read JWST data.
\texttt{stingray} has also had a major performance overhaul. The package
\texttt{gammapy} \cite{gammapy} also has major performance improvements,
has unified its API in preparation for its first stable release.

>>>>

The \astropypkg core package is mature and stable, in that many features have
been part of the core package for years with (purposefully) few changes to the
software interface.
This maturity and stability allows for the broader astronomy \python community
to rely upon the \astropypkg core package and build specialized packages within
the Astropy ecosystem.
Even many non-astronomical \python libraries have come to rely on \astropypkg.

The relationship between \astropy and community is reciprocal, with the growing and evolving needs of the community and the maturation of the broader scientific \python ecosystem driving the development of the \astropypkg core package since \citealt{astropy:2018}.
% feels like ↑ could be expanded to 2 sentences

% Maybe briefly outline the sectors of  
% Using \astropypkg for
% precision astronomy has spurred development of faster and more accurate coordinate transformations and time calculations,
% super-computer work has spurred \astropypkg to be thread-safe and support \package{dask} \citep{daskpkg}
% Astropy has grown to interoperate, e.g. NumPy

Development may be roughly split into a few categories: new features added, performance enhancements, intra and inter-package interoperability, and improvements to precision, accuracy, and reproducibility.

% Summarize major pieces of development since the last paper (since v2.0). A few
% ideas (taken from What's New pages) below, but feel free to expand on or remove
% from this list:
% \begin{itemize}
%     \item Support for representing and transforming velocity data in coordinates, epoch propagation (v3.0)
%     \item Improved support for astronomical time series: TimeSeries object (v3.2), Box Least Squares periodogram (v3.1)
%     \item Overall improved support of Quantity throughout numpy (v4.0) and scipy
%     \item Native support for Time, Quantity, and SkyCoord objects in Astropy tables
%     \item APE 14 WCS
%     \item A new SpectralCoord class for representing and transforming spectral quantities
%     \item Support for dask/large arrays in several places in the core package (4.1, 5.0, and maybe other versions)
% \end{itemize}

\subsubsection*{New Features} \label{sec:core-features-new}

\subsubsection*{Interoperability} \label{sec:core-features-interoperability}

\subsection*{Precision, Accuracy, & Reproducibility} \label{sec:core-features-precision_accuracy_repreducibility}

\subsection*{Precision, Accuracy, & Reproducibility} \label{sec:core-features-performance}

\subsection*{Precision, Accuracy, & Reproducibility} \label{sec:core-features-future}

@@
